---
title: "Introduction to divermeta"
author: "Felipe González-Casabianca, Vladimir Mikryukov"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to divermeta}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `divermeta` package provides tools for calculating various biodiversity metrics with a focus on multiplicity index. 
This index implements methods for quantifying diversity at different organizational levels, accounting for both abundance and dissimilarity between elements.

The core indexes implemented in this package include:

* **Inventory multiplicity**: The ratio of diversity before and after clustering elements.
* **Distance multiplicity**: A measure based on the ratio of functional diversities.

In addition, there are several other functions for estimation of previously proposed diversity indexes 
(e.g., Rao's quadratic entropy, Chiu & Chao 2014 functional diversity, etc.).

## Installation

You can install the development version of divermeta from GitHub with:

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("Mycology-Microbiology-Center/divermeta")
```

Let's load the package to explore its functionality:

```{r setup}
library(divermeta)
```

## Inventory multiplicity

Inventory multiplicity is defined as the ratio between the true diversity of unclustered abundances and the clustered abundances. 
It provides a way to quantify the diversity lost when grouping elements together.

### Basic example

Let's start with a simple example of calculating inventory multiplicity:

```{r}
# High multiplicity example
# Three clusters of ten different units each
# Expected multiplicity is close to 10
ab_high <- rep(10, 30)
clust_high <- c(rep(1, 10), rep(2, 10), rep(3, 10))
multiplicity.inventory(ab_high, clust_high)

# Low multiplicity example
# Three clusters with one unit each
# Expected multiplicity is close to 1
ab_low <- rep(10, 3)
clust_low <- c(1, 2, 3)
multiplicity.inventory(ab_low, clust_low)
```

In the high multiplicity example, we have 30 units divided into 3 clusters of 10 units each. 
The multiplicity is close to 10, indicating that each cluster contains approximately 10 distinct elements.

In the low multiplicity example, we have 3 units in 3 separate clusters, 
resulting in a multiplicity close to 1, indicating no information is lost when clustering.


